### 2017-05-18 Tesi-frn.tex
Add files via upload

%% This file has been automatically generated by `frontespizio'.
%% Don't use it as a model for a new frontispiece, use the
%% `frontespizio' environment in you document instead.
\documentclass [titlepage]{article}
\usepackage[utf8x]{inputenc}
\usepackage [a4paper,left=1cm,bottom=1.5cm,right=1cm,top=1cm]{geometry}
\usepackage {graphicx}
\usepackage [svgnames]{xcolor}
\usepackage[nowrite,infront,standard]{frontespizio}
\nofiles 
\fontoptionsans 
\Universita {Studi di Padova}
\Dipartimento {Fisica e Astronomia}
\Corso [Laurea Magistrale]{Fisica}
\Logo [1.5cm]{unipd}
\Annoaccademico {2016--2017}
\Titoletto {Tesi di laurea magistrale}
\Titolo {The mass dependence of dark matter halo alignments \\ with large-scale structure}
\Sottotitolo {}
\Candidato [1114287]{Davide Piras}
\Relatore {Prof.~Sabino Matarrese}
\NCorrelatore {Supervisor}{Supervisors}
\Correlatore {Dr Benjamin Joachimi \\ Dept. of Physics and Astronomy \\ UCL, United Kingdom \\[\frontrelcorrelsep ] \frontfixednamesfont Controrelatore\csname front@punct\endcsname \\ Prof.ssa Giulia Rodighiero}
\begin {document}
\preparefrontpagestandard
\end {document}